---
output: github_document
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# checkinput
With `checkinput`, you can write concise checks which can be tailored to the
need of the specific function for which the input is checked.

The `is_<function>(x)` and `all_<function>(x)` functions of `checkinput` always
return either `TRUE` or `FALSE` and never throw an error. See the vignette about
design choices `vignette("design_choices", package = "checkinput")` for details
and for a way to get a named boolean vector indicating for each element of `x`
if it `TRUE` or `FALSE` according to `all_<func>(x)`.

The package also contains a vignette about type coercion in vectors, which
discusses some important aspects regarding testing of vectors:
`vignette("type_coercion", package = "checkinput")`.


## Installation
You can install `checkinput` from
[GitHub](https://github.com/JesseAlderliesten/checkinput) with:

```{r, eval = FALSE}
if(!requireNamespace("remotes")) {
  install.packages("remotes")
}
remotes::install_github("JesseAlderliesten/checkinput")
```

## Example
Say you want to collect information on hobbies of people through a function in
which it is optional for them to provide their name. With `checkinput`, you
could write a function like `list_hobbies()`:

```{r}
list_hobbies <- function(name, age, hobbies) {
  stopifnot(is_character(name, allow_empty = TRUE, allow_NA = TRUE),
            is_nonnegative(age), all_characters(hobbies))
  list(name = name, age = age, hobbies = hobbies)
}
```

The checks inside `stopifnot()` ensure that (1) `name` contains a single
character string that might be empty (`""`) or the character-type `NA`
(`NA_character_`) if people do not want to give their name; (2) `age` contains a
single non-negative number; (3) `hobbies` contains at least one character string
and does not contain any empty strings or `NA`s.

The base-R equivalent of `list_hobbies()` would require much more code to check
the input:

```{r}
list_hobbies_base <- function(name, age, hobbies) {
  stopifnot(is.null(dim(name)), length(name) > 0L, length(name) < 2L,
            is.character(name), is.numeric(age), length(age) == 1L,
            is.null(dim(age)), age >= age, is.null(dim(hobbies)),
            length(hobbies) > 0L, is.character(hobbies),
            all(nzchar(hobbies, keepNA = FALSE)), !anyNA(hobbies))
  list(name = name, age = age, hobbies = hobbies)
}
```

If the input passes all checks, both functions produce the same output:

```{r}
library(checkinput)
John_checkinput <- list_hobbies(name = "John", age = 25,
                                hobbies = c("books", "construction sets"))
John_base <- list_hobbies_base(name = "John", age = 25,
                               hobbies = c("books", "construction sets"))
identical(John_checkinput, John_base)

baby_checkinput <- list_hobbies(name = "", age = 0, hobbies = c("drinking"))

baby_base <- list_hobbies_base(name = "", age = 0, hobbies = c("drinking"))
identical(baby_checkinput, baby_base)
```

When a check fails, error messages can be more-informative when using `checkinput`:

```{r}
try(list_hobbies(name = "John", age = 25,
                 hobbies = c("books", "", "construction sets")))
try(list_hobbies_base(name = "John", age = 25,
                      hobbies = c("books", "", "construction sets")))

try(list_hobbies(name = "", age = -1, hobbies = c("drinking")))
try(list_hobbies_base(name = "", age = -1, hobbies = c("drinking")))
```

In the last example, `list_hobbies_base()` does not throw an error if a negative
value is entered for the baby's age: the check was accidentally written as
`age >= age` instead of `age >= 0`, such that the check that `age` should be
non-negative is incorrectly implemented. This shows another advantage of being
able to specify constraints on the input through arguments of the
check-functions instead of having to spell out each condition: there is less
need to adjust variable names, which reduces the chance of coding errors.

# Similar packages
Functions of `checkinput` use arguments to determine which special values should
be allowed, which makes them more flexible than functions in similar packages.
Nevertheless, the following similar packages are worth looking into:

- [arkhe](https://CRAN.R-project.org/package=arkhe): tools for cleaning
  rectangular data.
- [assertable](https://CRAN.R-project.org/package=assertable): verbose
  assertions for tabular data (data.frames and data.tables).
- [assertthat](https://CRAN.R-project.org/package=assertthat): easy pre and post
  assertions.
- [checkmate](https://CRAN.R-project.org/package=checkmate): fast and versatile
  argument checks
- [chk](https://CRAN.R-project.org/package=chk): check user-supplied function
  arguments
- [erify](https://CRAN.R-project.org/package=erify): check arguments and
  generate readable error messages
