---
title: "Type coercion"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{type_coercion}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Introduction
Objects in R are converted ('coerced') to a common type when combined in a
vector (see `help("c")`). The type of the vector, and thus of all its
components, will be the
highest type of the components in the hierarchy `NULL` < `raw` < `logical` <
`integer` < `double` < `complex` < `character` < `list` < `expression`, see the
section `Details` in `help("c")` and the section `Value` in `help("typeof")`.
For example, numeric
`314` will be coerced to character `"314"` when it is combined in a vector with
character `"nco"`, such that `c(314, "nco")` results in the character vector
`c("314", "nco")`. This also holds for the logical `NA`, which will be coerced
to `NA_character_` that is printed as `NA`.

```{r Type coercion}
num <- 314
str(num)
char <- "nco"
str(char)
str(c(num, char))

logiNA <- NA
str(logiNA)
logiNA_char <- c(logiNA, char)
str(logiNA_char)
str(logiNA_char[1])
logiNA_char[1]
```


## Consequences for checks
As a consequence of type coercion, code like
`all_characters(c(x, y))` does **not** check if all elements in `x` and `y` have
type `character`: `x` and `y` will be coerced to the highest of their types
before the check is performed, such that `all_characters(c(x, y))` checks if any
of `x` or `y` has type character **and** none of `x` or `y` has a higher type.

To check if all elements in `x` and `y` have type `character`, use
`all_characters(x) && all_characters(y)` or, to generalise more easily to more
than two objects, `all(unlist(lapply(X = list(x, y), FUN = all_characters)))`.
Other function arguments passed to `all_characters()`, e.g., `allow_empty = TRUE`
to allow empty quotes (i.e., `""`), should be placed behind argument `x`.

Elements of a list can have different types but using `unlist()` on a list
creates a vector in which all elements of the list are coerced to the highest of
their types, such that `all_characters(unlist(z))` does **not** check if all
elements of list `z` are character but if any element of `z` is a character
**and** none of the elements of `z` is of a higher type.

```{r}
library(checkinput)

x <- 1:3
all_characters(x)

y <- letters[x]
all_characters(y)

all_characters(c(x, y)) # TRUE, even though 'x' is numeric!
all_characters(x) && all_characters(y)
all(unlist(lapply(X = list(x, y), FUN = all_characters)))

z <- list("a", 2, "c")
z
lapply(X = z, class)
# TRUE, even though the second element of 'z' is numeric!
all_characters(unlist(z))
all(unlist(lapply(X = z, FUN = all_characters)))
```


## Zero-length values
Although zero-length objects (see `help("is_zerolength")`) are discarded when
combined
into a vector with other values, their types are taken into account for type
coercion. For example, numeric `314` will be coerced to character `"314"` when
it is combined into a vector with zero-length `character(0)`, such that
`c(314, character(0))` results in the character string `"314"`, **not** in the
numeric value `314`.

```{r}
num <- 314
str(num)
zerochar <- character(0)
str(zerochar)
str(c(num, zerochar))
```
